Carrier phase and amplitude estimation using pilots and data

### 2012-05-05 paper.tex
Added a bit more to stats

In this setting, the transmitted symbols take the form,
\[
s_i = e^{j u_i},
\]
where $j = \sqrt{-1}$ and $u_i$ is from the set $\{0, \tfrac{2\pi}{M}, \dots, (M-1)\tfrac{2\pi}{M}\}$ and $M$ is the size of the constellation, i.e. we have assumed an $M$-PSK constellation.  We assume that some of the transmitted symbols are \emph{pilot symbols} known to the receiver and the remainder are information carrying \emph{data symbols} with phase that is unknown to the receiver.  So,
\[
s_i = \begin{cases}
p_i & i \in P, \\
d_i & i \in D, \\
0 & otherwise,
\end{cases}
\]
where $P$ is the set of indices describing the position of the pilot symbols $p_i$, and $D$ is a set of indices describing the position of the data symbols $d_i$.  The sets $P$ and $D$ are disjoint, i.e. $P \cap D = \emptyset$, and the union $P \cup D$ contains all those indices where the transmitter is active.  Outside $P \cup D$ the transmitter is silent (not active).

@@

We are interested in analysing $\hat{\theta}$ the maximiser of $\abs{Z(\theta)}^2$.  Put $\lambda_L = \langle\theta_0 - \hat{\theta}\rangle$.  We want to show that $\lambda_L$ converges to zero as the number of transmitted symbols $L$ increases to $\infty$ where $\sfracpart{x}$ denotes $x$ taken `modulo $2\pi$' into the interval $[-\pi, \pi)$, that is
\[
\fracpart{x} = x - 2\pi\operatorname{round}\left(\frac{x}{2\pi}\right).
\]
Under the assumption that the noise $\{w_i\}$ is circularly symetric then
\[
y_i = a_0 s_i + w_i = \rho_j e^{j ( \theta_0 + \Phi_i + k_i) }
\] 
where $k_i = \angle s_i$ and the $\Phi_i$ are identically distributed.  The $\Phi_i$ can be identified as \emph{circular random variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.  Recalling the definition of $\hat{d}_i(\theta)$ and $\hat{u}_i(\theta)$ from~\eqref{eq:hatdfinxtheta}, we have
\begin{align*}
\hat{u}_i(\theta) &= \round{\angle{y_i} - \theta} \\
&= \round{\theta_0 + \Phi_i + k_i - \theta} \\
&\equiv \round{ \fracpart{\theta_0 - \theta} + \Phi_i + k_i} \pmod M \\
&= \round{ \lambda + \Phi_i + k_i } \\
\end{align*}
where $\lambda = \fracpart{\theta_0 - \theta}$.  So,
\begin{align*}
 y_i e^{-j\theta} \hat{d}_i(\theta)^* &= \rho_i e^{j(\lambda + \Phi_i + k_i - \round{\lambda + \Phi_i + k_i})} \\
&= \rho_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})}
\end{align*}
when $i \in D$ and
\[
y_i e^{-j\theta} p_i^* = \rho_i e^{j(\lambda + \Phi_i)}
\]
when $i \in P$.  Now we can write
\[
Z(\theta) = \sum_{i \in P} \rho_i e^{j(\lambda + \Phi_i)} + \sum_{i \in D}  \rho_i e^{j( \lambda + \Phi_i - \round{\lambda + \Phi_i})}.
\]
Put $S_L(\lambda) = \frac{1}{L}\abs{Z(\theta)}^2$ and put $\lambda_L = \langle\theta_0 - \hat{\theta}\rangle$.  Then $\lambda_L$ is the maximiser of $S_L(\lambda)$.  We will show that $\lambda_L$ converges almost surely to zero as $L \rightarrow \infty$.  From this, the proof of strong consistency follows.

@@

We now prove the asymptotic normality of the estimator.  At the same time we are able to describe the distribution of the magnitude estimator $\hat{\rho}$.  Observe that $Z(\theta) = e^{-j\theta}Y(\{\hat{d}_i(\hat{\theta}), i \in D\})$.  From~\eqref{eq:ahatYhat} we obtain
\begin{align*}\label{eq:hata}
\hat{a} e^{-j\theta} &= \hat{\rho} e^{j(\hat{\theta} - \theta)} = \hat{\rho} e^{j\lambda_L} \\
&= \frac{1}{L} e^{-j\theta} Y(\{\hat{d}_i(\hat{\theta}), i \in D\}) \\
&= \frac{1}{L}  Z(\hat{\theta}) \\
&= \frac{1}{L} \sum_{i \in P} \rho_i e^{j(\lambda_L + \Phi_i)} + \frac{1}{L}  \sum_{i \in D}  \rho_i e^{j( \lambda_L + \Phi_i - \round{\lambda_L + \Phi_i})}
\end{align*}
from which it follows that
\begin{align*}\label{eq:hatrhoworking}
\hat{\rho} &=  \frac{1}{L} \sum_{i \in P} \rho_i e^{j\Phi_i} + \frac{1}{L}  \sum_{i \in D}  \rho_i e^{j(\Phi_i - \round{\lambda_L + \Phi_i})} \\
 &=  \frac{1}{L} \sum_{i \in P} (\rho_0 + w_i) + \frac{1}{L}  \sum_{i \in D}  (\rho_0 + w_i)e^{-j\round{\lambda_L + \Phi_i}} 
\end{align*}

>>>>

In this setting, the transmitted symbols take the form,
\[
s_i = e^{j u_i},
\]
where $j = \sqrt{-1}$ and $u_i$ is from the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$ and $M$ is the size of the constellation, i.e. we have assumed an $M$-PSK constellation.  We assume that some of the transmitted symbols are \emph{pilot symbols} known to the receiver and the remainder are information carrying \emph{data symbols} with phase that is unknown to the receiver.  So,
\[
s_i = \begin{cases}
p_i & i \in P, \\
d_i & i \in D, \\
0 & otherwise,
\end{cases}
\]
where $P$ is the set of indices describing the position of the pilot symbols $p_i$, and $D$ is a set of indices describing the position of the data symbols $d_i$.  The sets $P$ and $D$ are disjoint, i.e. $P \cap D = \emptyset$, and the union $P \cup D$ contains all those indices where the transmitter is active.  Outside $P \cup D$ the transmitter is silent (not active).

@@

We are interested in analysing $\hat{\theta}$ the maximiser of $\abs{Z(\theta)}^2$.  Put $\lambda_L = \langle\theta_0 - \hat{\theta}\rangle$.  We want to show that $\lambda_L$ converges to zero as the number of transmitted symbols $L$ increases to $\infty$ where $\sfracpart{x}$ denotes $x$ taken `modulo $2\pi$' into the interval $[-\pi, \pi)$, that is
\[
\fracpart{x} = x - 2\pi\operatorname{round}\left(\frac{x}{2\pi}\right).
\]
Under the assumption that the noise $\{w_i\}$ is circularly symetric then
\[
y_i = a_0 s_i + w_i = \rho_j e^{j ( \theta_0 + \Phi_i + k_i) }
\] 
where $k_i = \angle s_i$ and the $\Phi_i$ are identically distributed.  The $\Phi_i$ can be identified as \emph{circular random variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.  Recalling the definition of $\hat{d}_i(\theta)$ and $\hat{u}_i(\theta)$ from~\eqref{eq:hatdfinxtheta}, we have
\begin{align*}
\hat{u}_i(\theta) &= \round{\angle{y_i} - \theta} \\
&= \round{\theta_0 + \Phi_i + k_i - \theta} \\
&\equiv \round{ \fracpart{\theta_0 - \theta} + \Phi_i + k_i} \pmod 2\pi \\
&= \round{ \lambda + \Phi_i + k_i } \\
\end{align*}
where $\lambda = \fracpart{\theta_0 - \theta}$.  So,
\begin{align*}
 y_i e^{-j\theta} \hat{d}_i(\theta)^* &= \rho_i e^{j(\lambda + \Phi_i + k_i - \round{\lambda + \Phi_i + k_i})} \\
&= \rho_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})}
\end{align*}
when $i \in D$ and
\[
y_i e^{-j\theta} p_i^* = \rho_i e^{j(\lambda + \Phi_i)}
\]
when $i \in P$.  Now we can write
\[
Z(\theta) = \sum_{i \in P} \rho_i e^{j(\lambda + \Phi_i)} + \sum_{i \in D}  \rho_i e^{j( \lambda + \Phi_i - \round{\lambda + \Phi_i})}.
\]
Put $S_L(\lambda) = \frac{1}{L}\abs{Z(\theta)}^2$ and put $\lambda_L = \langle\theta_0 - \hat{\theta}\rangle$.  Then $\lambda_L$ is the maximiser of $S_L(\lambda)$.  We will show that $\lambda_L$ converges almost surely to zero as $L \rightarrow \infty$.  From this, the proof of strong consistency follows.

@@

We now prove the asymptotic normality of the estimator.  At the same time we are able to describe the distribution of the magnitude estimator $\hat{\rho}$.  Observe that $Z(\theta) = e^{-j\theta}Y(\{\hat{d}_i(\hat{\theta}), i \in D\})$.  From~\eqref{eq:ahatYhat} we obtain
\begin{align*}\label{eq:hata}
\hat{a} e^{-j\theta} &= \hat{\rho} e^{j(\hat{\theta} - \theta)} = \hat{\rho} e^{j\lambda_L} \\
&= \frac{1}{L} e^{-j\theta} Y(\{\hat{d}_i(\hat{\theta}), i \in D\}) \\
&= \frac{1}{L}  Z(\hat{\theta}) \\
&= \frac{1}{L} \sum_{i \in P} \rho_i e^{j(\lambda_L + \Phi_i)} + \frac{1}{L}  \sum_{i \in D}  \rho_i e^{j( \lambda_L + \Phi_i - \round{\lambda_L + \Phi_i})}
\end{align*}
and muliplying both sides of this equation by $e^{-j\lambda_L}$,
\begin{align*}\label{eq:hatrhoworking}
\hat{\rho} &=  \frac{1}{L} \sum_{i \in P} \rho_i e^{j\Phi_i} + \frac{1}{L}  \sum_{i \in D}  \rho_i e^{j(\Phi_i - \round{\lambda_L + \Phi_i})} \\
 &=  \frac{1}{L} \sum_{i \in P} (\rho_0 + w_i) + \frac{1}{L}  \sum_{i \in D}  (\rho_0 + w_i)e^{-j\round{\lambda_L + \Phi_i}} 
\end{align*}
